---
title: "DS311 - R Lab Assignment"
author: "Your Name"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    highlight: tango
    df_print: paged
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Assignment 1

* In this assignment, we are going to apply some of the build in data set in R for descriptive statistics analysis.
* To earn full grade in this assignment, students need to complete the coding tasks for each question to get the result.
* After finished all the questions, knit the document into HTML format for submission.

### Question 1

Using the **mtcars** data set in R, please answer the following questions.

```{r}
# Loading the data
data(mtcars)

# Head of the data set
head(mtcars)
```

a. Report the number of variables and observations in the data set.

```{r}
# Enter your code here
nrow(mtcars)
ncol(mtcars)

# Answer:
print("There are total of __11___ variables and __32___ observations in this data set.")
```



b. Print the summary statistics of the data set and report how many discrete and continuous variables are in the data set.

```{r}
# Enter your code here!
summary(mtcars)
library(magrittr)
mtcars %>% summary()

# Answer:
print("There are __5___ discrete variables and ___6__ continuous variables in this data set.")
```



c. Calculate the mean, variance, and standard deviation for the variable **mpg** and assign them into variable names m, v, and s. Report the results in the print statement.

```{r}
# Enter your code here!
m = mean(mtcars$mpg)
v = var(mtcars$mpg)
s = sd(mtcars$mpg)

print(paste("The average of Mile Per Gallon from this data set is ", m , " with variance ", v , " and standard deviation", s , "."))
```



d. Create two tables to summarize 1) average mpg for each cylinder class and 2) the standard deviation of mpg for each gear class.

```{r}
# Enter your code here!
install.packages("plyr")
install.packages("dplyr")
library(plyr)
library(dplyr)
cyl_avg_mpg =
  mtcars %>%
  group_by(cyl) %>%
  summarise(AvgMpg=mean(mpg))

gear_avg_mpg =
  mtcars %>%
  group_by(gear) %>%
  summarise(AvgMpg=mean(mpg))

cyl_avg_mpg
gear_avg_mpg


```



e. Create a crosstab that shows the number of observations belong to each cylinder and gear class combinations. The table should show how many observations given the car has 4 cylinders with 3 gears, 4 cylinders with 4 gears, etc. Report which combination is recorded in this data set and how many observations for this type of car.

```{r}
# Enter your code here!
observe_by_cyl_gear =
  mtcars %>%
  group_by(cyl, gear) %>%
  count()

observe_by_cyl_gear
print("The most common car type in this data set is car with _8___ cylinders and _3___ gears. There are total of __8_ cars belong to this specification in the data set.")
```


*** 

### Question 2

Use different visualization tools to summarize the data sets in this question.

a. Using the **PlantGrowth** data set, visualize and compare the weight of the plant in the three separated group. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your findings.

```{r}
# Load the data set
data("PlantGrowth")

# Head of the data set
head(PlantGrowth)

# Enter your code here!
install.packages("ggplot2")
library(ggplot2)
g_plantgrowth = ggplot(data=PlantGrowth) + geom_boxplot(aes(x=group, y=weight))
g_plantgrowth = g_plantgrowth + labs(title="PlantGrowth", x="group", y="weight")
g_plantgrowth
```

Result:

=> Report a paragraph to summarize your findings from the plot!
Based on the findings the best spread of data that does not differ far from the min, max, and range is trt2.


b. Using the **mtcars** data set, plot the histogram for the column **mpg** with 10 breaks. Give labels to the title, x-axis, and y-axis on the graph. Report the most observed mpg class from the data set.

```{r}
g_mtcars = ggplot(data=mtcars) + geom_histogram(aes(x=mpg), colour="black", fill="white", binwidth = 10)
g_mtcars = g_mtcars + labs(title = "Distribution of MPG", x="Miles Per Gallon", y="Counts")
g_mtcars


print("Most of the cars in this data set are in the class of ___20___ mile per gallon.")
```



c. Using the **USArrests** data set, create a pairs plot to display the correlations between the variables in the data set. Plot the scatter plot with **Murder** and **Assault**. Give labels to the title, x-axis, and y-axis on the graph. Write a paragraph to summarize your results from both plots.

```{r}
# Load the data set
data("USArrests")

# Head of the data set
head(USArrests)

# Enter your code here!
install.packages("ggrepel")
library(ggrepel)
g_usarrests = ggplot(USArrests, aes(x=Murder, y=Assault, label=rownames(USArrests))) + geom_point()
g_usarrests = g_usarrests + labs(title="Murder VS Assault", x="Murder", y="Assault")
g_usarrests = g_usarrests + geom_text_repel(size=3)
g_usarrests
```

Result:

=> Report a paragraph to summarize your findings from the plot!

Places with low amounts of both assault cases and murder cases within the United States include North Dakota, Vermont, Hawaii, Wisconsin, Iowa, Maine, and other states within quadrint one. The states with high cases of both assult and murder are Florida, Mississippi, Georgia, South Carolina, and Louisiana to name a few.

***

### Question 3

Download the housing data set from www.jaredlander.com and find out what explains the housing prices in New York City.

Note: Check your working directory to make sure that you can download the data into the data folder. 

```{r, echo=FALSE}
# Load and clean the housing data set 
download.file(url='https://www.jaredlander.com/data/housing.csv',
              destfile='data/housing.csv', mode='wb')
housingData <- read.csv('data/housing.csv')
housingData <- subset(housingData, 
                      select = c("Neighborhood", "Market.Value.per.SqFt", "Boro", "Year.Built"))
housingData <- na.omit(housingData)                   
```

a. Create your own descriptive statistics and aggregation tables to summarize the data set and find any meaningful results between different variables in the data set.

```{r}
# Head of the cleaned data set
head(housingData)

# Enter your code here!
summary(housingData)
count_neighborhood = housingData %>% group_by(Neighborhood) %>% count()
count_boro = housingData %>% group_by(Boro) %>% count()
count_builtyear = housingData %>% group_by(Year.Built) %>% count()
count_builtyear_boro_neighborhood = housingData %>% group_by(Year.Built, Boro, Neighborhood) %>% count()
count_neighborhood
count_boro
count_builtyear
count_builtyear_boro_neighborhood

```

b. Create multiple plots to demonstrates the correlations between different variables. Remember to label all axes and give title to each graph.

```{r}
# Enter your code here!
bulityear_MVPS_TRIBECA = housingData[housingData$Neighborhood %in% "TRIBECA",]
bulityear_MVPS_TRIBECA = ggplot(bulityear_MVPS_TRIBECA, aes(x=Year.Built, y=Market.Value.per.SqFt)) + geom_point()
bulityear_MVPS_TRIBECA = bulityear_MVPS_TRIBECA + geom_smooth(method=lm)
bulityear_MVPS_TRIBECA = bulityear_MVPS_TRIBECA + labs(title="Market Value Per Square Foot VS Built Year in Financial", x="Built Year", y="Market Value Per Square Feet")
bulityear_MVPS_TRIBECA


g_built_boro = ggplot(count_boro, aes(Boro, n)) + geom_bar(stat="identity")
g_built_boro = g_built_boro + labs(title="House Built in Each Boro", x="Boro", y="House")
g_built_boro
```

c. Write a summary about your findings from this exercise.

=> Enter your answer here!

It seems as though the market value per square footage has had a steady increase since 1850 well into 2000 with a few outliers amoung the trendline. Most of the houses within this timeline were built within the manhattan boro of New York, followed by Brooklyn, Queens, the Bronx, and Staten Island.
